---
title: "Text analysis"
description: |
  Glasgow Climate Pact and Paris Agreement
date: 2022-03-13
output:
  distill::distill_article:
    self_contained: false
---

# Overview

This report performs a text analysis of the Glasgow Climate Pact and the Paris Agreement. First, word counts are compiled and compared between the two. Then a sentiment analysis is completed and compared.

Data citations: United Nations Framework Convention on Climate Change (UNFCCC) (2015). The Paris Agreement. https://unfccc.int/process-and-meetings/the-paris-agreement/the-paris-agreement

United Nations Framework Convention on Climate Change (UNFCCC) (2021). Glasgow Climate Pact. https://unfccc.int/documents/310475

# Setup

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(pdftools)
library(patchwork)
library(tidytext)
```

# Text input

```{r}
glasgow <- pdf_text(here("data", "glasgow.pdf"))
paris <- pdf_text(here("data", "paris.pdf"))
```

# Text wrangling

```{r}
glasgow <- glasgow %>% 
  str_remove_all("Advance unedited version") %>% 
  str_remove_all(pattern = "[0-9]+")
paris <- paris %>% 
  str_remove_all(pattern = "[0-9]+")

glasgow_lines <- data.frame(glasgow) %>% 
  mutate(page = 1:n()) %>% 
  mutate(full_text = str_split(glasgow, pattern = "\\n")) %>% 
  unnest(full_text) %>% 
  mutate(full_text = str_squish(full_text)) %>% 
  slice(-(1:4))
paris_lines <- data.frame(paris) %>% 
  mutate(page = 1:n()) %>% 
  mutate(full_text = str_split(paris, pattern = "\\n")) %>% 
  unnest(full_text) %>% 
  mutate(full_text = str_squish(full_text)) %>% 
  slice(-(1:10))

glasgow_nonstop_words <- glasgow_lines %>% 
  unnest_tokens(word, full_text) %>% 
  select(-glasgow) %>% 
  anti_join(stop_words, by = "word")
paris_nonstop_words <- paris_lines %>% 
  unnest_tokens(word, full_text) %>% 
  select(-paris) %>% 
  anti_join(stop_words, by = "word")
```

# Word counts

```{r}
glasgow_nonstop_word_count <- glasgow_nonstop_words %>% 
  count(word) %>% 
  mutate(freq = n/sum(n))
paris_nonstop_word_count <- paris_nonstop_words %>% 
  count(word) %>% 
  mutate(freq = n/sum(n))

glasgow_top25 <- glasgow_nonstop_word_count %>% 
  arrange(-freq) %>% 
  slice(1:25) %>% 
  mutate(word = fct_reorder(word, freq))
paris_top25 <- paris_nonstop_word_count %>% 
  arrange(-freq) %>% 
  slice(1:25) %>% 
  mutate(word = fct_reorder(word, freq))

glasgow_plot <- ggplot(glasgow_top25, aes(y = word, x = freq)) +
  geom_col(aes(fill = "red", alpha = word)) +
  labs(y = "Word", x = "Proportion of non-stop words", title = "Glasgow") +
  scale_x_continuous(limits = c(0, 0.06)) +
  theme_minimal() +
  theme(legend.position = "none")
paris_plot <- ggplot(paris_top25, aes(y = word, x = freq)) +
  geom_col(aes(fill = "red", alpha = word)) +
  labs(y = "Word", x = "Proportion of non-stop words", title = "Paris") +
  scale_x_continuous(limits = c(0, 0.06)) +
  theme_minimal() +
  theme(axis.title.y = element_blank(), legend.position = "none")

glasgow_plot + paris_plot
```

**Figure 1.** Top 25 non-stop words in the Glasgow Climate Pact and Paris Agreement. Data: UNFCCC.

# Sentiment analysis

```{r}
glasgow_bing <- glasgow_nonstop_words %>% 
  inner_join(get_sentiments("bing"), by = "word") %>% 
  group_by(sentiment) %>% 
  summarize(count = n())
paris_bing <- paris_nonstop_words %>% 
  inner_join(get_sentiments("bing"), by = "word") %>% 
  group_by(sentiment) %>% 
  summarize(count = n())

bing_combined <- glasgow_bing %>% 
  inner_join(paris_bing, by = "sentiment", suffix = c("_glasgow", "_paris")) %>% 
  pivot_longer(cols = count_glasgow:count_paris, names_to = "document", values_to = "count")

bing_combined$document <- bing_combined$document %>% 
  str_replace_all("count_glasgow", "Glasgow") %>% 
  str_replace_all("count_paris", "Paris")

ggplot(bing_combined, aes(x = document, y = count)) +
  geom_col(aes(fill = sentiment), position = "dodge") +
  scale_fill_manual(values = c("red3", "green4")) +
  labs(x = "Document", y = "Word count", fill = "Sentiment") +
  theme_minimal()
```

**Figure 2.** Sentiment analysis of the Glasgow Climate Pact and Paris Agreement, using the "bing" lexicon. The Paris Climate Agreement used more positive words. Data: UNFCCC.