---
title: "Binary logistic regression"
description: |
  Florida palmetto species
date: 2022-03-13
output:
  distill::distill_article:
    self_contained: false
---


# Overview

This code will use binary logistic regression to test the feasibility of using variables plant height, canopy length, canopy width, and number of green leaves to classify whether a palmetto is species *Serenoa repens* or *Sabal etonia*.

Data citation: Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

# Setup

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(GGally)
library(patchwork)
library(broom)
library(kableExtra)
library(caret)
```

# Data input and wrangling

```{r}
palmetto <- read_csv(here("data", "palmetto.csv")) %>% 
  select(species, height, length, width, green_lvs) %>% 
  drop_na() %>% 
  mutate(species_name = case_when(species == 1 ~ "Serenoa repens",
                                  species == 2 ~ "Sabal etonia")) %>% 
  mutate(species_name = as.factor(species_name))
```

# Data visualization

```{r}
plot1 <- ggplot(palmetto, aes(x = height, y = length)) +
  geom_point(aes(color = species_name), size = 1, show.legend = FALSE) +
  labs(x = "Maximum height (cm)", y = "Maximum canopy length (cm)") +
  theme_minimal()

plot2 <- ggplot(palmetto, aes(x = height, y = width)) +
  geom_point(aes(color = species_name), size = 1) +
  labs(x = "Maximum height (cm)", y = "Maximum canopy width (cm)", color = "Species") +
  theme_minimal()

plot1 + plot2
```

**Figure 1.** Canopy length and width of Florida palmettos by height and species. It is somewhat easier to distinguish species based on canopy length than width.

```{r}
ggplot(palmetto, aes(x = species_name, y = green_lvs)) +
  geom_boxplot() +
  labs(x = "Species", y = "Number of green leaves") +
  theme_minimal()
```

**Figure 2.** *Serenoa repens* tends to have more green leaves than *Sabal etonia*.

# Binomial logistic regression and cross validation

```{r}
f1 <- species_name ~ height + length + width + green_lvs
f2 <- species_name ~ height + width + green_lvs

set.seed(42)

tr_ctrl <- trainControl(method = 'repeatedcv', repeats = 10, number = 10)

### train the model
cv_training_model1 <- train(f1, data = palmetto,
                method = "glm", family = "binomial",
                trControl = tr_ctrl)

cv_training_model2 <- train(f2, data = palmetto,
                method = "glm", family = "binomial",
                trControl = tr_ctrl)

blr_model1 <- glm(formula = f1, data = palmetto, family = "binomial")
blr_model2 <- glm(formula = f2, data = palmetto, family = "binomial")
```

Based on 10 repeats of 10-fold cross validation, the model that includes canopy length has a higher classification accuracy than the one without canopy length.

# Model results

```{r}
blr_model1_table <- tidy(blr_model1) %>% 
  select(-statistic) %>% 
  mutate(p.value = ifelse(p.value < 0.001, "< 0.001"))

colnames(blr_model1_table) <- c("Term", "Estimate of coefficient", "Standard error", "p value")
kable(blr_model1_table, caption = "Table 1. Parameters of binomial logistic regression model.") %>% 
  kable_styling("striped")
```


```{r}
blr_model1_prob <- augment(blr_model1, type.predict = "response") %>% 
  mutate(predicted_sp = case_when(.fitted > 0.5 ~ "Serenoa repens",
                                  .fitted < 0.5 ~ "Sabal etonia"))
blr_model1_grouped <- blr_model1_prob %>% 
  group_by(species_name, predicted_sp) %>% 
  summarize(count = n())

final_table <- tribble(~Species, ~correct, ~incorrect, ~percent,
                       "Sabal etonia", blr_model1_grouped$count[1], blr_model1_grouped$count[2], (blr_model1_grouped$count[1] / (blr_model1_grouped$count[1] + blr_model1_grouped$count[2])),
                       "Serenoa repens", blr_model1_grouped$count[4], blr_model1_grouped$count[3], (blr_model1_grouped$count[4] / (blr_model1_grouped$count[3] + blr_model1_grouped$count[4])))
colnames(final_table) <- c("Species", "# correctly classified", "# incorrectly classified", "% correctly classified")

kable(final_table, caption = "Table 2. Accuracy of binomial logistic regression model.") %>% 
  kable_styling("striped")
```